---
categories:
- ""
- ""
date: "2020-9-16"
description: Gender discrimination
draft: false
keywords: ""
slug: gendercase
title: A Case Study on Gender Discrimination
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(plotly)
```

## Background information

At the last board meeting of ABC Group, the headquarters of a large multinational company, the issue was raised that women were being discriminated in the company, in the sense that the salaries were not the same for male and female executives. A quick analysis of a sample of 50 employees (of which 24 men and 26 women) revealed that the average salary for men was about 8,700 higher than for women. This seemed like a considerable difference, so it was decided that a further analysis of the company salaries was warranted. In the following sections an analysis is carried out.

## Objective
The objective is to find out whether there is indeed a significant difference between the salaries of men and women, and whether the difference is due to discrimination or whether it is based on another, possibly valid, determining factor. 

## Data anlysis

### Loading the data


```{r load_omega_data}
genderpay <- read_csv(here::here("data", "gender.csv"))
glimpse(genderpay) # examine the data frame
```

### Determine the relationship between Salary and Gender

```{r, confint_single_valiables}
# Summary Statistics of salary by gender
mosaic::favstats (salary ~ gender, data=genderpay)

# Dataframe with two rows (male-female) and having as columns gender, mean, SD, sample size, 
# the t-critical value, the standard error, the margin of error, 
# and the low/high endpoints of a 95% condifence interval
salary_sum_stats <- genderpay %>%
  group_by(gender) %>%
  summarise(mean_salary = mean(salary),
            sd_salary = sd(salary),
            count = n(),
            se_salary = sd_salary / sqrt(count),
            t_critical = qt(0.975, count - 1),
            margin_of_error = t_critical * se_salary,
            salary_low = mean_salary - margin_of_error,
            salary_high = mean_salary + margin_of_error)

salary_sum_stats
```

#### Analysis:

In this sample size, male's average salary is higher than the female's. Male's 95% confidence interval is also higher than the female's with no overlaps. If we take another sample, we are quite confident that the male's average salary would be higher than the female's in that sample too.

```{r, hypothesis_testing1}
# hypothesis testing using t.test() 
t.test(salary ~ gender, data = genderpay)

# hypothesis testing using infer package
salary_diff <- genderpay %>%
  infer::specify(salary ~ gender) %>%
  infer::calculate(stat = "diff in means", order = c("male", "female"))

salary_gender_null <- genderpay %>%
  infer::specify(salary ~ gender) %>%
  infer::hypothesize(null = "independence") %>%
  infer::generate(reps = 1000, type = "permute") %>%
  infer::calculate(stat = "diff in means", order = c("male", "female"))

salary_gender_null %>% visualize() +
  infer::shade_p_value(obs_stat = salary_diff, direction = "two-sided")

salary_gender_null %>%
  infer::get_p_value(obs_stat = salary_diff, direction = "two_sided")
```

#### Anlysis:  
The null hypothesis is rejected. The mean salaries of male and female are significantly different.

### Explore the relationship between Professional Experience and Gender


```{r, experience_stats}
# Summary Statistics of salary by gender
favstats (experience ~ gender, data=genderpay)

```

### Hypothesis testing

```{r, hypothesis_testing}
# hypothesis testing using t.test() 
t.test(experience ~ gender, data = genderpay)

# hypothesis testing using infer package
experience_diff <- genderpay %>%
  specify(experience ~ gender) %>%
  calculate(stat = "diff in means", order = c("male", "female"))

experience_gender_null <- genderpay %>%
  specify(experience ~ gender) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("male", "female"))

experience_gender_null %>% visualize() +
  shade_p_value(obs_stat = experience_diff, direction = "two-sided")

experience_gender_null %>%
  get_p_value(obs_stat = experience_diff, direction = "two_sided")
```

#### Analysis:
There is a significant difference between the average experience of the male and female executives. But it does not prove the correlation between salary and experience, so it can neither validate nor endanger the previous conclusion.

### Explore the relationship between Professional Experience and Salary

```{r, salary_exp_scatter}
salary_exp <- ggplot(
  genderpay, 
  aes(
    x=experience, 
    y=salary
    )) +
  geom_point() +
  geom_smooth() +
  labs(
    title = "The Relationship between Salary and Experience",
    x = "Experience",
    y = "Salary"
  )
salary_exp <- ggplotly(salary_exp)
salary_exp


```


<br/>

### Check correlations between the data

```{r, ggpairs}
genderpay %>% 
  select(gender, experience, salary) %>% #order variables they will appear in ggpairs()
  ggpairs(aes(colour=gender, alpha = 0.3))+
  theme_bw()
```

#### Analysis:
Salary and experience have a strong correlation and the correlation is higher for female category than male. 

### Findings


